Challenges and Solutions of Developing and Implementing a Novel Desktop-as-a-Service

### 2024-05-04 paper.tex
Minor changes (missing spaces, remove comments, add TODO's)

% TODO: Christian (1)
% TODO: Johannes (2)

\section{Related Work}
\label{sec:relatedworkAbschnitt}
% TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
% TODO: Christian (1)
% TODO: Johannes (2)

@@

\begin{itemize}
	\item Transfer time (upstream) for sending the user input to the service provider.
	\item Time required for capturing the frames and encoding it as a video stream.
	\item Transfer time (downstream) for sending the video stream to the client.
	\item Time required for decoding the video stream back into a frame.
\end{itemize}

For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications. Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms. This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance. As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between the service provider and the consumer should not be greater than 1000 miles (approx. 1600\,km) because otherwise, the round-trip communications delay time between will be too long for video games.

@@

\section{Architecture}
\label{sec:RecommendDaaSarchitecture}
% 
% TODO: Christian (1)
% TODO: Johannes (2)

Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).

\begin{figure}
	\centering
	\includegraphics[width=.75\linewidth]{images/DaaS_DESIGN_Architecture_v12_english.pdf}
	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.} 
	\label{figure_architecture}
\end{figure}

@@

% TODO: Christian (1)
% TODO: Johannes (2)

@@

\subsection{Compatibility}
% 
% TODO: Christian (1)
% TODO: Johannes (2)

@@

\subsection{Performance}
% Initial brainstorm:
% + Welche Performance-Paramter / Kriterien sind hier relevant für die Anwendung / Nutzung
% + Fokus: Bottlenecks, Möglichkeiten der Performance-Steigerung, Grenzen/Limits, die man akzeptieren muss
% + Latenz hängt auch immer von Entfernung und Übertragungsmedium ab.
% + Skalierbarkeit: Stand und Möglichkeite
% + Ceph-Dateisystem
% + Datenbank SQlite aktuell. Perspektivisch MariaDB
% + Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host.
%
% Structure:
% performance of:
% - our serice itself
% - backend services
% - Viewer performance
% Limiting Factors
% - Hosting
% - Network
% - Internal data and communication
% Chosen techniques/solutions/protocols:
% - Filesystem (Ceph)
% - Database (sqlite, mariadb)
% - Guacamole (Proxy approach)
% Similar_Examples:
% - Cloud Gaming efforts 
%    - technical solutions, 
%    - latencies, 
%    - hardware support and integration)
%
% TODO: Johannes (1)
% TODO: Christian (2)
% TODO: Christian: Latenz bei Cloud-Gaming Erfahrungen
% TODO: 
%       - Include overall takeaway messages somewhere before:
%           - IaaS, PaaS and SaaS combined is defined as DaaS
%           - Define terms: Backend-service, Host-System, Instance
%           - We want a stable and usable system first, performance comes second
%           - Low effort to use and integrate is the most crucial aspect!!!
%           - No "Gamestation", but all modern applications shall be suported
%               - "Regular" office apps (Spreadsheets, Editor, Email, Filemanager)
%               - 3D modelling (Blender, Mayo, etc.)
%               - Audio/Video Editing? (kdenlive, Adobe Premiere, OBS?)
%               - Services (Webserver, Emailserver, Fileserver, etc.)
%           - If several solutions are plausible, implement all and let the user decide
%           - Give definitions for all specified daas-usecase
%       - More references
%       - Illustrate each subsection
%       - Typos, grammar, etc.

To describe the mission-critical performance aspects of the discussed system,
we initially provide a brief overview of potential challenges
by outlining bottlenecks and other relevant problem areas.
This is done in the context of the underlying use case model of DaaS systems.
Subsequently, we give a more comprehensive description
of our specific design decisions and tailored solutions.

@@

% From our point of view, the three most critical performance aspects
% for Daas-Systems are:
% \begin{itemize}
% 	\item Infrastructure and Platform
% 	\item Network Latency and Bandwidth
% 	\item Internal Processing and Communication
% \end{itemize}
% Since each of them has their own characteristics,
% specific behavior and requirements, implying a potential impact
% on the overall system performance and resource utilization,
% we describe them individually and in more detail
% in the following separate subsections.

@@

For SaaS, performance largely depends on the underlying usage model
and the actual software being virtualized.
Depending on the use case, influential properties
such as multi-tenancy, provision strategies, or the virtualized data type
might be utilized to varying degrees, resulting in overall
better or worse performing systems ~\cite{zhong2010virtualization}.

@@

\subsection{Stability}
% TODO: Johannes (1)
% TODO: Christian (2)
% TODO: Johannes: Beschreibung zur Position der Komponenten, Möglichkeiten der Parallelisierung und Einschränkungen
%
% Was ist hier relevant?
% Verfügbarkeit des Service und der Benutzerdaten
% Redundante Komponenten (Hardware und Software)
% Welche Komponenten kann man sinnvoll redundant vorhalten?
% Was sind die Vor- und Nachteile?
% Auch Kosten?
% Wo sehen wir aktuell in unserer Architektur Risiken für Dienstausfall und Datenverlust?
% Features von Proxmox
% Ceph-Dateisystem
% Datenbank SQlite aktuell. Perspektivisch MariaDB
% Guacamole läuft 1x direkt auf dem Host aktuell. Läuft theoretisch auf jedem Host. Proxmox kann sich darum kümmern?!
% Docker haben/nicht haben / Vor- und Nachteile
% Hier auch auf die Skalierbarkeit beziehen
% Structure:
%	- Stability
%		- Redundancy
%		- Scalability
%	- Backend-services
%		- Database
%		- Authentication
%		- Distributed Filesystem
%		- Container/VM (Proxmox, Kubernetes?) 
%		- Instance error => Recreate
%	- Considerations:
%		- TODO: Mention that term: Divide and conquer
%		- Requirement: Distributable components
%	- Risks:
%		- Filesystem error
%		- Database error
%		- Authentication error
%		- Infrastructure error (Container/VM)
%	- Benefits/Drawbacks:
%		- Distribution of components is distribution of load => Performance
%		- Eliminates bottlenecks
%		- Errors affect only errornous component

Stability in DaaS applications is especially important
for meeting the demands of continuous and time-critical operations
across diverse user scenarios.
Ensuring the stability of all system components during runtime is paramount
to maintain service continuity, handle growing user loads,
and transparently recover from failures
while providing a seamless end-user experience.
These considerations are integral to our architectural concept,
directly impacting the operational integrity of the system
and its suitability for enterprise-level deployment and trust.

@@

For Linux operating system deployments,
the free software projects xrdp and x11vnc enable all desired features
and single-app functionality can be easily obtained
by using a custom window manager such, as for example, qtile
~\footnote{\url{https://qtile.org/}}.

Table~\ref{tab:DaaS_Services_Overview} includes an overview
of the mentioned service implementations for RDP and VNC protocols
and their relevant characteristics.
Further implementations like TigerVNC and protocols like NX NoMachine exist
but do not seem useful for designing and implementing a novel DaaS because
they do not meet all feature requirements.~\cite{OJCC_2023v8i1n01_Baun}

@@

One challenge was imposed by Guacamole's current lack of support
for dynamic desktop resizing with VNC,
leading to misaligned layouts and a poor user experience.
To address this until such features are available in Guacamole
\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
we implemented a separate mechanism using a WebSocket proxy
and direct interaction with each instance.

@@

% \section{Outlook}
% \label{sec:NextSteps}

@@

% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{biblio}
%

% \bibliography{\jobname_no_dashes_at_repeated_names,references}{}

>>>>

\section{Related Work}
\label{sec:relatedworkAbschnitt}

@@

\begin{itemize}
	\item Transfer time (upstream) for sending the user input to the service provider.
	\item Time required for capturing the frames and encoding it as a video stream.
	\item Transfer time (downstream) for sending the video stream to the client.
	\item Time required for decoding the video stream back into a frame.
\end{itemize}
% TODO: Mention hardware-accelerated framebuffer protocols like \cite{hanimplementation} \cite{darcy2022latency}
% TODO: More scientific papers: \cite{di2021network} \cite{carrascosa2022cloud}
For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications. Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms. This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance. As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between the service provider and the consumer should not be greater than 1000 miles (approx. 1600\,km) because otherwise, the round-trip communications delay time between will be too long for video games.

@@

\section{Architecture}
\label{sec:RecommendDaaSarchitecture}
% 
Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).
\begin{figure}
	\centering
	\includegraphics[width=.75\linewidth]{images/DaaS_DESIGN_Architecture_v12_english.pdf}
	\caption{Architecture of the novel DaaS DESIGN~\cite{OJCC_2023v8i1n01_Baun}.}
	\label{figure_architecture}
\end{figure}

@@

@@

\subsection{Compatibility}

@@

\subsection{Performance}
To describe the mission-critical performance aspects of the discussed system,
we initially provide a brief overview of potential challenges
by outlining bottlenecks and other relevant problem areas.
This is done in the context of the underlying use case model of DaaS systems.
Subsequently, we give a more comprehensive description
of our specific design decisions and tailored solutions.

@@

@@

For SaaS, performance largely depends on the underlying usage model
and the actual software being virtualized.
Depending on the use case, influential properties
such as multi-tenancy, provision strategies, or the virtualized data type
might be utilized to varying degrees, resulting in overall
better or worse performing systems~\cite{zhong2010virtualization}.

@@

\subsection{Stability}
Stability in DaaS applications is especially important
for meeting the demands of continuous and time-critical operations
across diverse user scenarios.
Ensuring the stability of all system components during runtime is paramount
to maintain service continuity, handle growing user loads,
and transparently recover from failures
while providing a seamless end-user experience.
These considerations are integral to our architectural concept,
directly impacting the operational integrity of the system
and its suitability for enterprise-level deployment and trust.

@@

For Linux operating system deployments,
the free software projects xrdp and x11vnc enable all desired features
and single-app functionality can be easily obtained
by using a custom window manager such,
as for example, qtile~\footnote{\url{https://qtile.org/}}.

Table~\ref{tab:DaaS_Services_Overview} includes an overview
of the mentioned service implementations for RDP and VNC protocols
and their relevant characteristics.
Further implementations like TigerVNC and protocols like NX NoMachine exist
but do not seem useful for designing and implementing a novel DaaS because
they do not meet all feature requirements~\cite{OJCC_2023v8i1n01_Baun}.

@@

One challenge was imposed by Guacamole's current lack of support
for dynamic desktop resizing with VNC,
leading to misaligned layouts and a poor user experience.
To address this until such features are available in
Guacamole~\footnote{\url{https://issues.apache.org/jira/browse/GUACAMOLE-1196}},
we implemented a separate mechanism using a WebSocket proxy
and direct interaction with each instance.

@@

@@

% ---- Bibliography ----
\bibliographystyle{splncs04}
\bibliography{biblio}
